brados: declarative, programmable object storage

### 2016-05-25 paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

The Brados language corresponds to a subset of the Bloom language, and models the
storage system state uniformly as a collection of
relations~\cite{alvaro:cidr11}. The composition of a collection of existing
interfaces is then expressed as a collection of high-level \emph{queries} that
describe how a stream of requests (API calls) are filtered, transformed and
combined with existing state to define streams of outputs (API call returns as
well as updates to system state).  Separating the relational semantics of such
compositions from details of their physical implementations introduces a
number of degrees of freedom, similar to the ``data independence’’ offered by
database query languages.  The choice of access methods (for example, whether
to use a bytestream interface or a key/value store), storage device
classes (e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
1:1 or N:1 mapping) and execution plans (e.g. operator ordering) can be
postponed to execution time.  The optimal choices for these physical design
details are likely to change much more often than the logical specification,
freeing the interface designer from the need to rewrite systems and device and
interface characteristics change.

@@

We now show the implementation of CORFU using Brados and use this as a vehicle
for describing additional aspects of the language of semantics.
Listing~\ref{lst:state} shows the declaration of state for the CORFU
interface. Lines 1 and 2 define the schema of the two persistent collections
that hold the current epoch value, and the log contents. These collections are
mapped onto storage within Ceph but abstract away the low-level interface
(e.g. bytestream vs key-value). Lines 5-9 define the input and output
interfaces. We use a generic schema for the input operation to simplify
how rules are defined that apply to all operation types.

@@

  scratch :write_op, [:type, :pos, :epoch] => [:data]
  scratch :read_op,  [:type, :pos, :epoch] => [:data]
  scratch :trim_op,  [:type, :pos, :epoch] => [:data]
  scratch :fill_op,  [:type, :pos, :epoch] => [:data]
  scratch :seal_op,  [:type, :pos, :epoch] => [:data]

  # op did or did not pass the epoch guard
  scratch :valid_op,   [:type, :pos, :epoch] => [:data]
  scratch :invalid_op, [:type, :pos, :epoch] => [:data]

  # op's position was or was not found in the log
  scratch :found_op,    [:type, :pos, :epoch] => [:data]
  scratch :notfound_op, [:type, :pos, :epoch] => [:data]
end
\end{lstlisting}

\begin{lstlisting}[caption={Setup}, label=lst:setup]
bloom do
  # op's position found in log
  found_op <= (valid_op * log).lefts(pos => pos)
  notfound_op <= valid_op.notin(found_op)

@@

Listing~\ref{lst:eg} shows the epoch guard that is applied to all operations.
The guard rejects requests that are tagged with old epoch values, ensuring that
a client generating a request has an up-to-date view of the system.  First the
\texttt{invalid\_op} collection is defined to include the current operation if
its epoch value is no larger than the stored epoch value. Next the
\texttt{valid\_op} collection is defined to be the inverse of
\texttt{invalid\_op} and is a helper used to refine other operations later in
the dataflow. Finally we handle the case for all operations tagged with an
out-of-date epoch by merging the \texttt{invalid\_op} set into the output
\texttt{ret} collection.

\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
bloom :epoch_guard do
  invalid_op <= (op * epoch).pairs{|o,e|
    o.epoch <= e.epoch}
  valid_op <= op.notin(invalid_op)
  ret <= invalid_op{|o|
    [o.type, o.pos, o.epoch, 'stale']}
end
\end{lstlisting}

The process of sealing an object requires installing a new epoch value and
returning the current maximum position written. Listing~\ref{lst:seal} 
implements the seal interface by first removing the current epoch value and
replacing it with the epoch value contained in the operation. Next an aggregate
is computed over the log to find the maximum position written, and this value
is returned.

@@

Trimming a log entry always succeeds. In Listing~\ref{lst:trim} The
\texttt{<+-} operator simulatenously removes the log entry with the given 
position and replaces it with an entry with its state set to \emph{trimmed}.

@@

The write and fill interfaces are implemented similarily, and are show in
Listing~\ref{lst:write} and Listing~\ref{lst:fill}, respectively.  A valid
write collection is created if the operation position is not found in the log.
A valid write operation is then merged into log, otherwise a read only error is
returend indicating that hte log position was already written to.
The fill operation is identical except a the fill state is set on the log entry.

\begin{lstlisting}[caption={Write}, label=lst:write]
bloom :write do
  temp :valid_write <= write_op.notin(found_op)
  log <+ valid_write{ |o| [o.pos, 'valid', o.data]}
  ret <= valid_write{ |o|
    [o.type, o.pos, o.epoch, 'ok'] }
  ret <= write_op.notin(valid_write) {|o|
    [o.type, o.pos, o.epoch, 'read-only'] }
end
\end{lstlisting}

\begin{lstlisting}[caption={Fill}, label=lst:fill]
bloom :fill do
  temp :valid_fill <= fill_op.notin(found_op)
  log <+ valid_fill { |o| [o.pos, 'fill'] }
  ret <= valid_fill { |o|
    [o.type, o.pos, o.epoch, 'ok'] }
  ret <= fill_op.notin(valid_fill) { |o|
    [o.type, o.pos, o.epoch, 'read-only'] }
end
\end{lstlisting}

The read interface is shown in Listing~\ref{lst:read}, and structured similar
to the write interface. First we create a collection containing a valid read
operation that is in the log and does not have the filled or trimmed state set.
The data is returend in the case of a valid read op, otherwise an error is
returend through the output interface.

>>>>

The Brados language corresponds to a subset of the Bloom language, and models the
storage system state uniformly as a collection of
relations~\cite{alvaro:cidr11}. The composition of a collection of existing
interfaces is then expressed as a collection of high-level \emph{queries} that
describe how a stream of requests (API calls) are filtered, transformed and
combined with existing state to define streams of outputs (API call returns as
well as updates to system state).  Separating the relational semantics of such
compositions from details of their physical implementations introduces a
number of degrees of freedom, similar to the ``data independence'' offered by
database query languages.  The choice of access methods (for example, whether
to use a bytestream interface or a key/value store), storage device
classes (e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
1:1 or N:1 mapping) and execution plans (e.g. operator ordering) can be
postponed to execution time.  The optimal choices for these physical design
details are likely to change much more often than the logical specification,
freeing the interface designer from the need to rewrite systems and device and
interface characteristics change.

@@

We now show the implementation of CORFU using Brados and use this as a vehicle
for describing additional aspects of the language of semantics.
Listing~\ref{lst:state} shows the declaration of state for the CORFU
interface. Lines 1 and 2 define the schema of the two persistent collections
that hold the current epoch value, and the log contents. These collections are
mapped onto storage within Ceph but abstract away the low-level interface
(e.g. bytestream vs key-value). Lines 5-9 define the input and output
interfaces. We use a generic schema for the input operation to simplify
how rules are defined that apply to all operation types. Lines 11-15 define
named collections for each operation type. The scratch type indicates that the
data is not persistent, and only remains in the collection for a single 
execution time step. The remaining scratch collections are defined to further
subdivide the operations based on different properties which we'll describe next.

@@

  scratch :write_op, op.schema
  scratch :read_op,  op.schema
  scratch :trim_op,  op.schema
  scratch :fill_op,  op.schema
  scratch :seal_op,  op.schema

  # op did or did not pass the epoch guard
  scratch :valid_op,   op.schema
  scratch :invalid_op, op.schema

  # op's position was or was not found in the log
  scratch :found_op,    op.schema
  scratch :notfound_op, op.schema
end
\end{lstlisting}

Initialization is performed in Listing~\ref{lst:setup} which acts as a demux
for the operation type and properties.  Lines 3-7 show the epoch guard that is
applied to all operations.  The guard rejects requests that are tagged with old
epoch values, ensuring that a client generating a request has an up-to-date
view of the system.  First the \texttt{invalid\_op} collection is defined to
include the current operation if its epoch value is no larger than the stored
epoch value. Next the \texttt{valid\_op} collection is defined to be the
inverse of \texttt{invalid\_op} and is a helper used to refine other operations
later in the dataflow. Finally we handle the case for all operations tagged
with an out-of-date epoch by merging the \texttt{invalid\_op} set into the
output \texttt{ret} collection.

Lines 10 and 11 populate the \texttt{found\_op} and \texttt{notfound\_op}
collections that allow operations behavior to be predicated on if the position
associated with a request is found in the log. Finally the remaining lines in
Listing~\ref{lst:setup} populate each of the specific operation collections.

\begin{lstlisting}[caption={Setup}, label=lst:setup]
bloom do
  # epoch guard
  invalid_op <= (op * epoch).pairs{|o,e|
    o.epoch <= e.epoch}
  valid_op <= op.notin(invalid_op)
  ret <= invalid_op{|o|
    [o.type, o.pos, o.epoch, 'stale']}

  # op's position found in log
  found_op <= (valid_op * log).lefts(pos => pos)
  notfound_op <= valid_op.notin(found_op)

@@

The process of sealing an object requires installing a new epoch value and
returning the current maximum position written. Listing~\ref{lst:seal}
implements the seal interface by first removing the current epoch value and
replacing it with the epoch value contained in the input operation. Next an
aggregate is computed over the log to find the maximum position written, and
this value is returned, typically to a client performing a reconfiguration of
the system or following the failure of a sequencer.

@@

Trimming a log entry always succeeds. In Listing~\ref{lst:trim} the
\texttt{<+-} operator simulatenously removes the log entry with the given 
position and replaces it with an entry with its state set to \emph{trimmed}.
In practice the removal of a log entry may trigger garbage collection, but
we model it here as an update for brevity.

@@

The write and fill interfaces are implemented similarily, and are both shown in
Listing~\ref{lst:write}.  A \texttt{valid\_write} collection is created if the
operation position is not found in the log.  A \texttt{valid\_write} operation
is then merged into log, otherwise a read only error is returend indicating
that the log position was already written to.  The fill operation is identical
except a the fill state is set on the log entry.

\begin{lstlisting}[caption={Write and Fill}, label=lst:write]
bloom :write do
  temp :valid_write <= write_op.notin(found_op)
  log <+ valid_write{ |o| [o.pos, 'valid', o.data]}
  ret <= valid_write{ |o|
    [o.type, o.pos, o.epoch, 'ok'] }
  ret <= write_op.notin(valid_write) {|o|
    [o.type, o.pos, o.epoch, 'read-only'] }
end

bloom :fill do
  temp :valid_fill <= fill_op.notin(found_op)
  log <+ valid_fill { |o| [o.pos, 'fill'] }
  ret <= valid_fill { |o|
    [o.type, o.pos, o.epoch, 'ok'] }
  ret <= fill_op.notin(valid_fill) { |o|
    [o.type, o.pos, o.epoch, 'read-only'] }
end
\end{lstlisting}

Finally the read interface is shown in Listing~\ref{lst:read}, and structured
in a similar way to the write and fill interfaces. First we create a collection
containing a valid read operation (named \texttt{ok\_read}) that is in the log
and does not have the filled or trimmed state set.  The data read from the log
is returend in the case of a valid read operation, otherwise an error is
returend through the output interface.
